Dirac neutrino mass generation from Majorana messenger

### 2019-09-18 draft.tex
Updates from Overleaf
mc

The most general Lagrangian for some gauge $\operatorname{U}(1)_X$ symmetry, which includes $X=\mathcal{D}$, contains the following gauge, Yukawa and scalar interactions:

@@

@@

with $V(\omega) = \mu^{2}_{\omega} \omega^{\dagger} \omega + \lambda_{\omega} (\omega^{\dagger} \omega)^{2}$. 

@@

Since we are interested here in keeping the exotic set of $X$-charges
charges $\{\mp 4$,~$\mp 4$,~$\pm 5\}$ in such a way that when assigned
to the right-handed neutrinos the tree level Dirac and Majorana masses
can be forbidden~\cite{Calle:2018ovc}, the set of Majorana mediators
$\psi_{R\alpha}$ which realize the dimension-6 operator at one-loop
have vector-like $X$-charges would spoil the anomaly cancellation
condition.
In view of that $\psi_{R\alpha}$ have necessarily nonzero $X$-charges (see the Appendix), we further add an extra set of chiral fermions in 
such a way the full set heavy fermions  do not affect
the anomaly cancellation (their charges cancel each other in a vector like way). 
The resulting charge assignment is displayed in
Table~\ref{tab:partcont3}, where 
the fields $\xi_{L \alpha}$ constitute the new set of chiral fermion that guarantee the anomaly cancellation.
%\footnote{Their Majorana masses are obtained in the same way than the ones for $\psi_{R\alpha}$.}.
In this way we end up with a model with four Majorana fermions, two more than in the minimal solution, since third generation of chiral fermions is not needed to cancel the $\operatorname{U(1)}_X$ anomalies (alternatively we may consider another simple setup by adding a single set of $\psi_R$ and $\xi_L$ and two set of scalars $\eta_\alpha$, $\sigma_{\alpha}$~\cite{Reig:2018mdk})\footnote{
%On the other hand, 
Seeing that the two additional left-handed fields $\xi_{L\alpha}$,  with a $X$-charge $r=3/4$,  have both Dirac and Majorana mass terms $\overline{\xi_{L \alpha}}\psi_{R \beta}$ and $ \overline{\xi_{L\alpha}^c }\xi_{L \beta} \langle S\rangle$, the full set of Majorana fields are massive and heavy. 
Moreover, because the $\operatorname{U}(1)_{X}$ left out a remnant $Z_2$ discrete symmetry which guarantees the stability of the lightest $Z_2$-odd particle,  we have that the lightest of these Majorana fields may play the role of DM candidate.}.

@@

\begin{figure}
    \centering
    \includegraphics[scale=1.0]{u1blc}
    \caption{Collider and cosmological constraints for the $\operatorname{U}(1)_{B-L}$ model.}
    \label{fig:u1blc}
\end{figure}

The constraint of $\Delta N_{\text{eff}}$ for $\operatorname{U}(1)_{B-L}$ is shown in the blue (middle) region and start to be better than current ATLAS limit for $M_{Z'}\gtrsim 4.8\ \text{TeV}$. Similar restrictions can be obtained for the other models quoted in Fig.~\ref{fig:Neff}.

@@

\section*{Acknowledgments}
Work supported by Sostenibilidad-UdeA and the UdeA/CODI Grant 2017-16286, and by COLCIENCIAS through the Grant 111577657253. 
\appendix
\section{$\operatorname{U}(1)_X$ anomaly cancellation }
We use $f$ ($f$) to denote the general $\operatorname{U}(1)_X$ generation-independent charge assignments of the field $f_R$ ($F_L$). The three linear anomalies in $\operatorname{U(1)_X}$~\cite{Campos:2017dgc}
\begin{align}
  \label{eq:anolin}
  \left[\mathrm{SU}(3)_{C}\right]^{2} \mathrm{U}(1)_{X} :& &
[3 u+3 d]-[3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{SU}(2)_{L}\right]^{2} \mathrm{U}(1)_{X} :&&
-[2 l+3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{U}(1)_{Y}\right]^{2} \mathrm{U}(1)_{X} :&&
{ \left[(-2)^{2} e+3\left(\tfrac{4}{3}\right)^{2} u+3\left(-\tfrac{2}{3}\right)^{2} d\right]-\left[2(-1)^{2} l+3 \cdot 2\left(\tfrac{1}{3}\right)^{2} q\right]}=&
 0,
  \end{align}
allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $N$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \left[\text{Grav}\right]^{2} \mathrm{U}(1)_{X}:\ \sum_{\alpha=1}^N n_{\alpha}+3 (e-2l)=&0\,, &  \left[\mathrm{U}(1)_{X}\right]^{3}:\ \sum_{\alpha=1}^N n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
We choose the solutions with $r\equiv e-2l$, such that
\begin{align}
  \label{eq:anolam}
  \sum_{\alpha=1}^{N} n_{\alpha}=&-3 r\,,   & \sum_{\alpha=1}^{N} n_{\alpha}^3=&-3 r^3\,.
\end{align}
The full set of standard model charges in terms of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc} that we choose as $l$ and $r$, is just
\begin{align}
  u=&-r-\frac{4l}{3}\,,&d=&r+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&r+2l\,,&h=&-r-l\,.
  \label{Eq:SMCharges}
\end{align}
If we define the Abelian symmetry in which only the right-handed charged fermions have non-vanishing $X$-charges as $\operatorname{U}(1)_R$, then the general anomaly free two-parameter solution can be written as
\begin{align}
  X=r R- l Y\,.
\end{align}

>>>>

The most general Lagrangian for some gauge $\operatorname{U}(1)_X$ symmetry, which includes the trivial case $X=\mathcal{D}$, must contains the following gauge, Yukawa and scalar interactions in
order to realize $\mathcal{O}_{6D}$ at one-loop:

@@

  \label{eq:VHSetasigma}

@@

with $V(\omega) = \mu^{2}_{\omega} \omega^{\dagger} \omega + \lambda_{\omega} (\omega^{\dagger} \omega)^{2}$. It is worth to emphasize that $\operatorname{U}(1)_{\mathcal{D}}$ automatically allows for all the terms in eqs.~\eqref{eq:LagY} and~\eqref{eq:VHSetasigma} but other realizations will be checked later on.

@@

Since we are interested here in keeping the exotic set of $X$-charges
charges $\{\mp 4$,~$\mp 4$,~$\pm 5\}$ in such a way that when assigned
to the right-handed neutrinos the tree level Dirac and Majorana masses
can be forbidden~\cite{Calle:2018ovc}, the set of Majorana mediators
$\psi_{R\alpha}$ which realize the dimension-6 operator at one-loop
 would spoil the anomaly cancellation
condition.
In view of that $\psi_{R\alpha}$ have necessarily nonzero $X$-charges (see the Appendix), we further add an extra set of chiral fermions in 
such a way the full set heavy fermions  do not affect
the anomaly cancellation (their charges cancel each other in a vector like way). 
The resulting charge assignment is displayed in
Table~\ref{tab:partcont3}, where 
the fields $\xi_{L \alpha}$ constitute the new set of chiral fermion that guarantee the anomaly cancellation.
%\footnote{Their Majorana masses are obtained in the same way than the ones for $\psi_{R\alpha}$.}.
In this way we end up with a model with four Majorana fermions, two more than in the minimal solution, since third generation of chiral fermions is not needed to cancel the $\operatorname{U(1)}_X$ anomalies (alternatively we may consider another simple setup by adding a single set of $\psi_R$ and $\xi_L$ and two set of scalars $\eta_\alpha$, $\sigma_{\alpha}$~\cite{Reig:2018mdk})\footnote{
%On the other hand, 
Seeing that the two additional left-handed fields $\xi_{L\alpha}$,  with a $X$-charge $r=3/4$,  have both Dirac and Majorana mass terms $\overline{\xi_{L \alpha}}\psi_{R \beta}$ and $ \overline{\xi_{L\alpha}^c }\xi_{L \beta} \langle S\rangle$, the full set of Majorana fields are massive and heavy. 
Moreover, because the $\operatorname{U}(1)_{X}$ left out a remnant $Z_2$ discrete symmetry which guarantees the stability of the lightest $Z_2$-odd particle,  we have that the lightest of these Majorana fields may play the role of DM candidate.}.

@@

\begin{figure}
    \centering
    \includegraphics[scale=1.0]{u1blc}
    \caption{Collider and cosmological constraints for the $\operatorname{U}(1)_{B-L}$ model, and contours of constant $g'$ in the plane of $q_s v_S$ as a function of $M_{Z'}$ with $q_S=2$.}
    \label{fig:u1blc}
\end{figure}

The constraint of $\Delta N_{\text{eff}}$ for $\operatorname{U}(1)_{B-L}$ is shown in the blue (middle) region and start to be better than current ATLAS limit for $M_{Z'}\gtrsim 4.8\ \text{TeV}$. In the figure we also show contours of constant $g'$ at $0.01$, $0.1$ and $0.5$ with the solid, dashed and dot-dashed lines respectively. We can see that start to constraint $g'$ for values larger than $0.1$ in $\operatorname{U}(1)_{B-L}$.    Similar restrictions can be obtained for the other models quoted in Fig.~\ref{fig:Neff}.

@@

\section*{Acknowledgments}
Work supported by Sostenibilidad-UdeA and the UdeA/CODI Grant 2017-16286, and by COLCIENCIAS through the Grant 111577657253. 
\appendix
\section{$\operatorname{U}(1)_X$ anomaly cancellation }
We use $f$ ($f$) to denote the general $\operatorname{U}(1)_X$ generation-independent charge assignments of the field $f_R$ ($F_L$). The three linear anomalies in $\operatorname{U(1)_X}$~\cite{Campos:2017dgc}
\begin{align}
  \label{eq:anolin}
  \left[\mathrm{SU}(3)_{C}\right]^{2} \mathrm{U}(1)_{X} :& &
[3 u+3 d]-[3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{SU}(2)_{L}\right]^{2} \mathrm{U}(1)_{X} :&&
-[2 l+3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{U}(1)_{Y}\right]^{2} \mathrm{U}(1)_{X} :&&
{ \left[(-2)^{2} e+3\left(\tfrac{4}{3}\right)^{2} u+3\left(-\tfrac{2}{3}\right)^{2} d\right]-\left[2(-1)^{2} l+3 \cdot 2\left(\tfrac{1}{3}\right)^{2} q\right]}=&
 0,
  \end{align}
allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $N$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \left[\text{Grav}\right]^{2} \mathrm{U}(1)_{X}:\ \sum_{\alpha=1}^N n_{\alpha}+3 (e-2l)=&0\,, &  \left[\mathrm{U}(1)_{X}\right]^{3}:\ \sum_{\alpha=1}^N n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
We choose the solutions with $r\equiv e-2l$, such that
\begin{align}
  \label{eq:anolam}
  \sum_{\alpha=1}^{N} n_{\alpha}=&-3 r\,,   & \sum_{\alpha=1}^{N} n_{\alpha}^3=&-3 r^3\,.
\end{align}
The full set of anomaly free standard model $X$-charges in terms of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc} that we choose as $l$ and $r$, is just
\begin{align}
  u=&-r-\frac{4l}{3}\,,&d=&r+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&r+2l\,,&h=&-r-l\,.
  \label{Eq:SMCharges}
\end{align}
By setting $l=0$ in the previous equations, we can define the Abelian symmetry in which only the right-handed charged fermions have non-vanishing $X$-charges as $\operatorname{U}(1)_R$. Then the general anomaly free two-parameter solution can be written as
\begin{align}
  X=r R- l Y\,.
\end{align}
